---
title: "Identifying various forms of conserved genes using cliques (master file)"
author: "Elena van Zalen"
date: "`r Sys.Date()`"
output: 
  html_document:
    code_folding: hide
    toc: true
    number_sections: true
    toc_float: true
    theme: spacelab
editor_options:
  chunk_output_type: console
---

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(igraph)
```

**JUMP TO STEP 3 FOR IDENTIFICATION OF VARIOUS CLIQUES**

**Step 1: remove orthogroups with no cliques**
PRODUCES: CONS_OGS_RED which contains all orthogroups with at least one clique (3-6 members)
```{r eval=FALSE, include=FALSE}
load("data/ComPlEx/all_expressed_genes.RData") 
expressologs <- expr_genes %>%
  rename(MaxpVal = Max.p.Val) %>% 
  filter(MaxpVal < 0.9)

load("data/ComPlEx/ones_and_zeros_all.RData")

# Add summary columns and remove all orthogroups that no expressed gene pairs.
df_with_sums <- ones_and_zeros_all %>% 
  mutate(Angiosperms = rowSums(. [1:6])) %>% 
  mutate(Cross = rowSums(. [7:14]))%>%  
  mutate(Gymnosperms = rowSums(. [15])) %>% 
  mutate(Conserved = rowSums(. [1:15])) %>% 
  filter(Conserved > 0)

cons_ogs <- df_with_sums%>% 
  filter(rownames(df_with_sums) %in% expressologs$OrthoGroup) %>% 
  rownames()

# Loop over all orthogroups from above and count the number of max cliques that contain at least three members.

numb_of_cliques_per_og <- c()

for (g in cons_ogs) {
  print(g)
  # g <- "OG0001993"
  
  expressologs_g <- expressologs %>% 
    filter(OrthoGroup == g) 
  
  expressologs_g <- expressologs_g %>% 
    mutate(UGeneSpecies1 = paste0(expressologs_g$Species1, "-", expressologs_g$GeneSpecies1),
           UGeneSpecies2 = paste0(expressologs_g$Species2, "-", expressologs_g$GeneSpecies2))
  
  nodes <- data.frame(name = unique(c(expressologs_g$UGeneSpecies1, expressologs_g$UGeneSpecies2)))
  
  edges <- data.frame(from = expressologs_g$UGeneSpecies1, 
                      to = expressologs_g$UGeneSpecies2)
  
  net <- graph_from_data_frame(edges, directed = FALSE, vertices = nodes)
  
  # largest_clique <- largest_weighted_cliques(net)
  largest_clique <- max_cliques(net, min = 3)

  numb_of_cliques_per_og <- rbind(numb_of_cliques_per_og, data.frame(Orthogroup = g,
                                                                     numbCliques = length(largest_clique),
                                                                     minSize = min(lengths(largest_clique)),
                                                                     maxSize = max(lengths(largest_clique))))
  }

# 2) Remove orthogroups with no cliques (min clique size = 3). The loop simply goes through all orthogroups and counts the number of cliques that contain at least three members. We therefore need to remove all orthogroups with zero cliques.
numb_of_cliques_per_og <- numb_of_cliques_per_og %>% 
  arrange(desc(numbCliques)) %>% 
  filter(numbCliques > 0) 

paste0("Number of orthologs with at least one clique (3 to 6 members): ", length(unique(numb_of_cliques_per_og$Orthogroup)))

# 3) Vectorise and save! 
cons_og_red <- numb_of_cliques_per_og$Orthogroup
save(cons_og_red, file = "data/ComPlEx/conserved_orthogroups_reduced.RData")
save(numb_of_cliques_per_og, file = "data/ComPlEx/number_of_cliques_per_OG.RData")

```

**Step 2: The clique code**
PRODUCES: clique_genes_prefiltered (if saved) and clique_genes_filterable. The latter is the final file which can be used for identifying all clique types.
```{r eval=FALSE, include=FALSE}
# load("Data/Data/number_of_cliques_per_OG.RData") # consider dropping OG0000001
#load("data/ComPlEx/conserved_orthogroups_reduced.RData") # orthogroups that contain at least one clique (size 3-6)
expressologs_from_max_cliques <- list()
max_numb_of_cliques <- 5500

for (g in cons_og_red) {
  print(g)
  
  # g <- "OG0000001"
  expressologs_g <- expressologs %>% 
    filter(OrthologGroup == g) 
  
  expressologs_g <- expressologs_g %>% 
    mutate(UGeneSpecies1 = paste0(expressologs_g$Species1, "-", expressologs_g$GeneSpecies1),
           UGeneSpecies2 = paste0(expressologs_g$Species2, "-", expressologs_g$GeneSpecies2))
  
  nodes <- data.frame(name = unique(c(expressologs_g$UGeneSpecies1, expressologs_g$UGeneSpecies2)))
  
  edges <- data.frame(from = expressologs_g$UGeneSpecies1, 
                      to = expressologs_g$UGeneSpecies2)
  
  net <- graph_from_data_frame(edges, directed = FALSE, vertices = nodes)
  
  largest_clique <- max_cliques(net, min = 3)

  if(length(largest_clique) > max_numb_of_cliques){
    largest_clique_red <- largest_clique[1:max_numb_of_cliques]
  } else{largest_clique_red <- largest_clique}
  
  for(c in 1:length(largest_clique_red)){
    c <- 1
    if (c %% 500 == 0) {
      cat(c, "\n")
    }
    
    clique <- largest_clique[[c]]
    clique_name <- attr(clique, "names")
    
    clique_species <- str_split_fixed(clique_name, "-", n = 2)[,1]
    clique_genes <-  str_split_fixed(clique_name, "-", n = 2)[,2]
    
    expressologs_in_clique <- expressologs_g %>%
      filter(UGeneSpecies1 %in% clique_name & UGeneSpecies2 %in% clique_name) %>%
      select(c(1:5, 8)) %>% 
      mutate(cliqueID = paste0(g,"_" ,c))
    
    expressologs_from_max_cliques <- append(expressologs_from_max_cliques, list(expressologs_in_clique))
  }}

clique_genes_unlisted <- plyr::ldply(expressologs_from_max_cliques)

clique_genes_unlisted_prefilter <- clique_genes_unlisted %>% 
  group_by(cliqueID) %>% 
  mutate(MaxpValNegLog10 = -log10(MaxpVal)) %>% 
  mutate(NegLog10CliqueSum = sum(MaxpValNegLog10)) %>% 
  ungroup()

# ------------- SAVING POINT -----------------
# As the code above does take some time to run - it may be a good idea to save progress here.
save(clique_genes_unlisted_prefilter, file = "data/ComPlEx/clique_genes_prefiltered.RData")
# load("data/ComPlEx/clique_genes_prefiltered.RData")

weighted_max_cliques <- clique_genes_unlisted_prefilter %>% 
  mutate(species_pair = paste0(Species1, Species2))

speciesPairs <- tibble(SpeciesPair = weighted_max_cliques %>% distinct(species_pair) %>% pull(species_pair),
                       SpeciesPairClade = c("Cross", "Cross", "Angio", "Angio", "Angio", "Cross", "Cross", "Cross",
                       "Cross", "Gymno", "Gymno", "Gymno", "Cross", "Cross", "Cross"))


weighted_max_cliques <-  left_join(weighted_max_cliques , speciesPairs, by = join_by("species_pair" == "SpeciesPair" ))

weighted_max_cliques_filterable <- weighted_max_cliques%>% 
  mutate(Clade = paste0(SpeciesPairClade, "-sum" ))%>% 
  group_by(cliqueID)%>% 
  mutate(OriginalCliqueSize = n())%>%  
  mutate(OriginalCliqueSize = factor(OriginalCliqueSize))%>% 
  ungroup() %>% 
  pivot_wider(names_from = SpeciesPairClade, values_from = SpeciesPairClade)%>%
  pivot_wider(names_from = Clade, values_from = MaxpValNegLog10)

weighted_max_cliques_filterable <- replace_na(weighted_max_cliques_filterable, list(`Angio-sum` = 0, `Gymno-sum` = 0, `Cross-sum` =0))

weighted_max_cliques_filterable <- weighted_max_cliques_filterable %>% 
  group_by(cliqueID) %>% 
  mutate(AngioSum = sum(`Angio-sum`)) %>% 
  mutate(GymnoSum = sum(`Gymno-sum`)) %>% 
  mutate(CrossSum = sum(`Cross-sum`))

weighted_max_cliques_filterable$Angio[is.na(weighted_max_cliques_filterable$Angio)] <- 0
weighted_max_cliques_filterable$Angio[weighted_max_cliques_filterable$Angio != 0] <-1
weighted_max_cliques_filterable$Gymno[is.na(weighted_max_cliques_filterable$Gymno)] <-0
weighted_max_cliques_filterable$Gymno[weighted_max_cliques_filterable$Gymno != 0] <-1
weighted_max_cliques_filterable$Cross[is.na(weighted_max_cliques_filterable$Cross)] <-0
weighted_max_cliques_filterable$Cross[weighted_max_cliques_filterable$Cross != 0] <-1

weighted_max_cliques_filterable <- weighted_max_cliques_filterable %>% 
  ungroup() %>% 
   mutate_at(c("Angio", "Gymno", "Cross"), as.numeric) %>% 
    mutate(UGeneSpecies1 = paste0(weighted_max_cliques_filterable$Species1, "-", weighted_max_cliques_filterable$GeneSpecies1),
           UGeneSpecies2 = paste0(weighted_max_cliques_filterable$Species2, "-", weighted_max_cliques_filterable$GeneSpecies2), .before = species_pair)

# ------------- SAVING POINT -----------------
# At this point the data set is ready to be used for identification of cliques.
save(weighted_max_cliques_filterable, file = "data/ComPlEx/cliques/clique_genes_filterable_coex005.RData")

```

**Step 3: identifying cliques**
Note that the two previous steps can be skipped if already run (and hopefully saved).
```{r}
# deposit ships this object as an .RDS; load() its .RData equivalent when regenerating
weighted_max_cliques_filterable <- readRDS("data/ComPlEx/cliques/clique_genes_filterable_COMPLETE.RDS")
load("data/ComPlEx/ones_and_zeros_all.RData")

```



```{r}
# --------------- DECISIONS ---------------
SELECTION_METHOD <- "SUM" # options: "SUM" = clique sum OR "PVAL" = p-value
TYPE_OF_GENE_SET <- "e" # options: "c" = conserved, "d" = differentiated, "e" = enhanced OR "u" = unique

# --------------- PARAMETERS ---------------
P_VAL <- 0.1 # P_VAL < x
C_SUM <- 15 # C_SUM >= x
CROSSED_PAIRS_PERMITTED <- 3 # CROSSED_PAIRS_PERMITTED <= x
C_SUM_CLADES <- 3 # C_SUM_CLADES >= x
C_SUM_CROSS <- 6 # C_SUM_CROSS <= x
C_SUM_TARGET_CLADE <- 3 # C_SUM_TARGET_CLADES >= x
C_SUM_NON_TARGET_CLADE <- 3 # C_SUM_MON_TARGET_CLADES < x

# NB! When identifying enhanced genes - an additional parameter has been added (04.09.24) removing orthogroups that only have 3-membered cliques. This is to reduce the possibility of overlapping with unique genes.


```

**Which parameters?**

Conserved
- p-value: P_VAL
- clique sum: C_SUM

Differentiated
- p-value: P_VAL + CROSSED_PAIRS_PERMITTED
- clique sum: C_SUM_CLADES + C_SUM_CROSS

Enhanced
- p-value: P_VAL
- clique sum: C_SUM_TARGET_CLADE + C_SUM_NON_TARGET_CLADE + C_SUM_CROSS

Unique
- p-value: P_VAL
- clique sum: C_SUM_TARGET_CLADE


```{r}
# ----------------CONSERVED GENES------------------------

if(TYPE_OF_GENE_SET == "c"){
  print(TYPE_OF_GENE_SET)

  if(SELECTION_METHOD == "PVAL"){
    print(SELECTION_METHOD)

    conserved_genes <- weighted_max_cliques_filterable %>%
      filter(MaxpVal < P_VAL) %>%
      group_by(cliqueID) %>%
      mutate(cliqueSize = n()) %>%
      filter(cliqueSize == 15)

    print(length(unique(conserved_genes$OrthologGroup)))
    file_name <- paste0("data/ComPlEx/conserved_genes_", SELECTION_METHOD, "_", P_VAL, ".RData")
  }

  if(SELECTION_METHOD == "SUM"){
    print(SELECTION_METHOD)
    conserved_genes <- weighted_max_cliques_filterable %>%
      group_by(cliqueID) %>%
      filter(NegLog10CliqueSum >= C_SUM) %>%
      mutate(cliqueSize = n()) %>%
      filter(cliqueSize == 15)

    print(length(unique(conserved_genes$OrthologGroup)))
    file_name <- paste0("data/ComPlEx/conserved_genes_", SELECTION_METHOD, "_", C_SUM, ".RData")
  }}

# ------------------DIFFERENTIATED GENES----------------------

if(TYPE_OF_GENE_SET == "d"){
  print(TYPE_OF_GENE_SET)

  if(SELECTION_METHOD == "PVAL"){
    print(SELECTION_METHOD)


    conserved_genes <- weighted_max_cliques_filterable %>%
      filter(MaxpVal < P_VAL) %>%
      group_by(cliqueID) %>%
      filter(sum(Angio) == 3 & sum(Gymno) == 3 & sum(Cross) <= CROSSED_PAIRS_PERMITTED) %>%
      filter(OriginalCliqueSize == 15)


    print(length(unique(conserved_genes$OrthologGroup)))
    file_name <- paste0("data/ComPlEx/differentiated_genes_", SELECTION_METHOD, "_", P_VAL,"_XPAIRS_",CROSSED_PAIRS_PERMITTED ,".RData")
  }

  if(SELECTION_METHOD == "SUM"){
    print(SELECTION_METHOD)

    conserved_genes <- weighted_max_cliques_filterable %>%
      group_by(cliqueID) %>%
      filter(AngioSum >= C_SUM_CLADES & GymnoSum >= C_SUM_CLADES & CrossSum < C_SUM_CROSS) %>%
      mutate(cliqueSize = n()) %>%
      filter(cliqueSize == 15)

    print(length(unique(conserved_genes$OrthologGroup)))
    file_name <- paste0("data/ComPlEx/differentiated_genes_", SELECTION_METHOD, "_C-X_", C_SUM_CLADES,"-",C_SUM_CROSS, ".RData")
  }}

# -----------------ENHANCED GENES-----------------------

if(TYPE_OF_GENE_SET == "e"){
  print(TYPE_OF_GENE_SET)

  if(SELECTION_METHOD == "PVAL"){
    print(SELECTION_METHOD)


    conserved_genes_angio <- weighted_max_cliques_filterable %>%
      filter(MaxpVal < P_VAL) %>%
      filter(!(OriginalCliqueSize == 3)) %>%  # Only orthogroups with cliques larger than 3.
      group_by(cliqueID) %>%
      filter(sum(Angio) == 3) %>%
      filter(sum(Gymno) == 0 & sum(Cross) == 0)

    print(length(unique(conserved_genes_angio$OrthologGroup)))

    conserved_genes_gymno <- weighted_max_cliques_filterable %>%
      filter(MaxpVal < P_VAL) %>%
      filter(!(OriginalCliqueSize == 3)) %>% # Only orthogroups with cliques larger than 3.
      group_by(cliqueID) %>%
      filter(sum(Gymno) == 3) %>%
      filter(sum(Angio) == 0 & sum(Cross) == 0)

    print(length(unique(conserved_genes_gymno$OrthologGroup)))


    file_name_angio <- paste0("data/ComPlEx/enhanced_angiosperm_genes_", SELECTION_METHOD, "_", P_VAL, ".RData")
    file_name_gymno <- paste0("data/ComPlEx/enhanced_gymnosperm_genes_", SELECTION_METHOD, "_", P_VAL, ".RData")

  }

  if(SELECTION_METHOD == "SUM"){
    print(SELECTION_METHOD)

    conserved_genes_angio <- weighted_max_cliques_filterable %>%
      filter(AngioSum >= C_SUM_TARGET_CLADE) %>%
      filter(GymnoSum < C_SUM_NON_TARGET_CLADE & CrossSum < C_SUM_CROSS) %>%
      group_by(cliqueID) %>%
      filter(sum(Angio) == 3) %>%
      mutate(cliqueSize = n()) %>%
      filter(cliqueSize > 3) # Removes orthogroups with only 3-membered cliques.

    print(length(unique(conserved_genes_angio$OrthologGroup)))

    conserved_genes_gymno <- weighted_max_cliques_filterable %>%
      filter(GymnoSum >= C_SUM_TARGET_CLADE) %>%
      filter(AngioSum < C_SUM_NON_TARGET_CLADE & CrossSum < C_SUM_CROSS) %>%
      group_by(cliqueID) %>%
      filter(sum(Gymno) == 3) %>%
      mutate(cliqueSize = n()) %>%
      filter(cliqueSize > 3) # Removes orthogroups with only 3-membered cliques.

    print(length(unique(conserved_genes_gymno$OrthologGroup)))
    file_name_angio <- paste0("data/ComPlEx/enhanced_angiosperm_genes_", SELECTION_METHOD, "_T-NT-X_", C_SUM_TARGET_CLADE,"-",C_SUM_NON_TARGET_CLADE, "-", C_SUM_CROSS,".RData")
    file_name_gymno <- paste0("data/ComPlEx/enhanced_gymnosperm_genes_", SELECTION_METHOD, "_T-NT-X_", C_SUM_TARGET_CLADE,"-",C_SUM_NON_TARGET_CLADE, "-", C_SUM_CROSS, ".RData")
  }

}

# ------------------UNIQUE GENES----------------------
if(TYPE_OF_GENE_SET == "u"){
  print(TYPE_OF_GENE_SET)

  if(SELECTION_METHOD == "PVAL"){
    print(SELECTION_METHOD)

    angio_unique_ogs <- ones_and_zeros_all %>%
      mutate(Angiosperms = rowSums(. [1:3])) %>%
      mutate(Cross = rowSums(. [4:12]))%>%
      mutate(Gymnosperms = rowSums(. [13:15])) %>%
      mutate(Conserved = rowSums(. [1:15])) %>%
      filter(Angiosperms == 3 & Gymnosperms == 0 & Cross == 0)

    angio_unique_ogs_vctr <- rownames(angio_unique_ogs)

    conserved_genes_angio <- weighted_max_cliques_filterable %>%
      filter(OrthologGroup %in% angio_unique_ogs_vctr) %>%
      filter(MaxpVal < P_VAL) %>%
      group_by(cliqueID) %>%
      filter(sum(Angio) == 3)


    print(length(unique(conserved_genes_angio$OrthologGroup)))


    gymno_unique_ogs <- ones_and_zeros_all %>%
      mutate(Angiosperms = rowSums(. [1:3])) %>%
      mutate(Cross = rowSums(. [4:12]))%>%
      mutate(Gymnosperms = rowSums(. [13:15])) %>%
      mutate(Conserved = rowSums(. [1:15])) %>%
      filter(Angiosperms == 0 & Gymnosperms == 3 & Cross == 0)

    gymno_unique_ogs_vctr <- rownames(gymno_unique_ogs)


    conserved_genes_gymno <- weighted_max_cliques_filterable %>%
      filter(OrthologGroup %in% gymno_unique_ogs_vctr) %>%
      filter(MaxpVal < P_VAL) %>%
      group_by(cliqueID) %>%
      filter(sum(Gymno) == 3)
    print(length(unique(conserved_genes_gymno$OrthologGroup)))


    file_name_angio <- paste0("data/ComPlEx/unique_angiosperm_genes_", SELECTION_METHOD, "_", P_VAL, ".RData")
    file_name_gymno <- paste0("data/ComPlEx/unique_gymnosperm_genes_", SELECTION_METHOD, "_", P_VAL, ".RData")

  }

  if(SELECTION_METHOD == "SUM"){
    print(SELECTION_METHOD)

    angio_unique_ogs <- ones_and_zeros_all %>%
      mutate(Angiosperms = rowSums(. [1:3])) %>%
      mutate(Cross = rowSums(. [4:12]))%>%
      mutate(Gymnosperms = rowSums(. [13:15])) %>%
      mutate(Conserved = rowSums(. [1:15])) %>%
      filter(Angiosperms == 3 & Gymnosperms == 0 & Cross == 0)

    angio_unique_ogs_vctr <- rownames(angio_unique_ogs)


    conserved_genes_angio <- weighted_max_cliques_filterable %>%
      filter(OrthologGroup %in% angio_unique_ogs_vctr) %>%
      filter(AngioSum >= C_SUM_TARGET_CLADE)

    print(length(unique(conserved_genes_angio$OrthologGroup)))


    gymno_unique_ogs <- ones_and_zeros_all %>%
      mutate(Angiosperms = rowSums(. [1:3])) %>%
      mutate(Cross = rowSums(. [4:12]))%>%
      mutate(Gymnosperms = rowSums(. [13:15])) %>%
      mutate(Conserved = rowSums(. [1:15])) %>%
      filter(Angiosperms == 0 & Gymnosperms == 3 & Cross == 0)

    gymno_unique_ogs_vctr <- rownames(gymno_unique_ogs)


    conserved_genes_gymno <- weighted_max_cliques_filterable %>%
      filter(OrthologGroup %in% gymno_unique_ogs_vctr) %>%
      filter(GymnoSum >= C_SUM_TARGET_CLADE)

    print(length(unique(conserved_genes_gymno$OrthologGroup)))

    file_name_angio <- paste0("data/ComPlEx/unique_angiosperm_genes_", SELECTION_METHOD,"_T_", C_SUM_TARGET_CLADE ,".RData")
    file_name_gymno <- paste0("data/ComPlEx/unique_gymnosperm_genes_", SELECTION_METHOD, "_T_", C_SUM_TARGET_CLADE ,".RData")
  }

}


```


SAVE

```{r}

# save(conserved_genes, file = file_name)
# save(conserved_genes_angio, file = file_name_angio)
# save(conserved_genes_gymno, file = file_name_gymno)

```