---
title: "Investigation of clique-genes (master file 4)"
author: "ELena van Zalen & Ellen Dimmen Chapple"
date: "`r Sys.Date()`"
output: 
  html_document:
    code_folding: hide
    toc: true
    number_sections: true
    toc_float: true
    theme: lumen
editor_options:
  chunk_output_type: console
---

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(cowplot)

# Ortholog group file
ORTHOLOG_GROUP_FILE <- read.delim("data/ComPlEx/Orthogroups_20240613.tsv", header = TRUE, sep = "\t")
```

# Conserved genes and differentiated genes
```{r}
# Gene set 
# path <- "data/ComPlEx/"
# file <- "conserved_genes_SUM_15.RData" 

#"differentiated_genes_SUM_C-X_3-9.RData"
#"conserved_genes_SUM_15.RData"
load("data/ComPlEx/cliques/conserved_genes_SUM_15coexpr005_full.RData")

# load(paste0(path, file))

gene_set <- conserved_genes %>% # One clique per orthogroup
  ungroup() %>% 
  group_by(OrthoGroup) %>% 
  arrange(desc(NegLog10CliqueSum)) %>% 
  slice(1:15)%>% 
  ungroup()

s1 <- gene_set %>% 
  select(GeneSpecies1) %>% 
  filter(grepl("Potra", GeneSpecies1)) %>% 
  rename("Genes" = GeneSpecies1)
  
s2 <- gene_set %>% 
  select(GeneSpecies2) %>% 
  filter(grepl("Potra", GeneSpecies2)) %>% 
  rename("Genes" = GeneSpecies2)

gene_set_aspen <- rbind(s1,s2)
gene_set_aspen <- gene_set_aspen %>% 
  distinct(Genes)

#gene_set_file_name <- paste0(path,"aspen_set_",file)
save(gene_set_aspen, file = "data/ComPlEx/geneset_aspen.RData")

# Expressed genes w/ annotation
expressed_genes <- read.table("data/expression/aspen_root_cold_stress_expression.txt", sep = "\t", row.names = 1, header = TRUE)
go_list_file <- read_delim("data/Potra22_blast2go_GO.txt",'\t')

go_list <- go_list_file %>% 
  separate(`Sequence Name`,into = c("Sequence Name","Sequence Variant"), sep = "\\.") %>%
  separate(`Annotation GO ID-Annotation GO Term`,into = c("GO ID","GO Term"), sep = "-", extra = "drop") %>%
  filter(`Sequence Name` %in% rownames(expressed_genes))

#save(go_list, file = "data/go_list_aspen.RData")

go_data_frame <- data.frame(go_id=go_list$`GO ID`, evidence=rep("ND", nrow(go_list)), 
                            gene_id=go_list$`Sequence Name`)

library(GSEABase)
library(GOstats)
library(gdata)

go_frame_object = GOFrame(go_data_frame,organism="Aspen")
go_all_frame_object = GOAllFrame(go_frame_object)

gene_set_collection <- GeneSetCollection(go_all_frame_object, setType = GOCollection())
universe <- rownames(expressed_genes)

gene_ID <- gene_set_aspen$Genes
group_name <- "conserved_genes"
pval_cutoff <- 0.05
params <- GSEAGOHyperGParams(name="GSEA",
                             geneSetCollection = gene_set_collection,
                             geneIds = gene_ID,
                             universeGeneIds = universe, 
                             ontology = "BP",
                             pvalueCutoff = pval_cutoff,
                             conditional = FALSE,
                             testDirection = "over")

hyper_g_test <- hyperGTest(params)
summary_table <- summary(hyper_g_test)
summary_table <- summary_table[summary_table$Count>1,]
summary_table <- summary_table[,c(1,2,5,6,7)]
colnames(summary_table) <- c("GO id","P-value","x","n","GO term")

rownames(summary_table) <- 1:nrow(summary_table)

summary_table$`P-value`<- format(summary_table$`P-value`, digits=3, scientific=TRUE)

# Save DON'T OVERWRITE
save(summary_table, file = "data/ComPlEx/SUMMARY_TBL_clique_GOterms.RData")
```

# Enhance and unique genes (angiosperm)
```{r}
# Gene set 
path <- "data/ComPlEx/"
file <- "enhanced_angiosperm_genes_SUM_T-NT-X_3-3-9.RData" 
# "enhanced_angiosperm_genes_SUM_T-NT-X_3-3-9.RData"
# "unique_angiosperm_genes_SUM_T_3.RData"

load(paste0(path, file))
gene_set <- conserved_genes_angio %>% # One clique per orthogroup
  ungroup() %>% 
  group_by(OrthoGroup) %>% 
  arrange(desc(NegLog10CliqueSum)) %>% 
  slice(1:15)%>% 
  ungroup()

s1 <- gene_set %>% 
  select(GeneSpecies1) %>% 
  filter(grepl("Potra", GeneSpecies1)) %>% 
  rename("Genes" = GeneSpecies1)
  
s2 <- gene_set %>% 
  select(GeneSpecies2) %>% 
  filter(grepl("Potra", GeneSpecies2)) %>% 
  rename("Genes" = GeneSpecies2)

gene_set_aspen <- rbind(s1,s2)
gene_set_aspen <- gene_set_aspen %>% 
  distinct(Genes)

# Expressed genes w/ annotation
expressed_genes <- read_delim("data/ComPlEx/transcriptomicsData/AspWood_transcriptomics.txt")
go_list_file <- read_delim("data/ComPlEx/Potra22_blast2go_GO.txt",'\t')

go_list <- go_list_file %>% 
  separate(`Sequence Name`,into = c("Sequence Name","Sequence Variant"), sep = "\\.") %>%
  separate(`Annotation GO ID-Annotation GO Term`,into = c("GO ID","GO Term"), sep = "-", extra = "drop") %>%
  filter(`Sequence Name` %in% expressed_genes$Genes)

go_data_frame <- data.frame(go_id=go_list$`GO ID`, evidence=rep("ND", nrow(go_list)), 
                            gene_id=go_list$`Sequence Name`)

library(GSEABase)
library(GOstats)
library(gdata)

go_frame_object = GOFrame(go_data_frame,organism="Aspen")
go_all_frame_object = GOAllFrame(go_frame_object)


gene_set_collection <- GeneSetCollection(go_all_frame_object, setType = GOCollection())
universe <- expressed_genes$Genes

gene_ID <- gene_set_aspen$Genes
group_name <- "conserved_genes"
pval_cutoff <- 0.05
params <- GSEAGOHyperGParams(name="GSEA",
                             geneSetCollection = gene_set_collection,
                             geneIds = gene_ID,
                             universeGeneIds = universe, 
                             ontology = "BP",
                             pvalueCutoff = pval_cutoff,
                             conditional = FALSE,
                             testDirection = "over")

hyper_g_test <- hyperGTest(params)
summary_table <- summary(hyper_g_test)
summary_table <- summary_table[summary_table$Count>1,]
summary_table <- summary_table[,c(1,2,5,6,7)]
colnames(summary_table) <- c("GO id","P-value","x","n","GO term")

rownames(summary_table) <- 1:nrow(summary_table)

summary_table$`P-value`<- format(summary_table$`P-value`, digits=3, scientific=TRUE)

# Save
file_for_saving <- paste0(path,"SUMMARY_TBL_", file)
save(summary_table, file = file_for_saving)
```

# Enhanced and unique genes - gymnosperms 
```{r}
library(tidyverse)

# Gene set 
path <- "data/ComPlEx/"
file <- "unique_gymnosperm_genes_SUM_T_3.RData" 

# "enhanced_gymnosperm_genes_SUM_T-NT-X_3-3-9.RData"
# "unique_gymnosperm_genes_SUM_T_3.RData"

load(paste0(path, file))
gene_set <- conserved_genes_gymno %>% # One clique per orthogroup
  ungroup() %>% 
  group_by(OrthoGroup) %>% 
  arrange(desc(NegLog10CliqueSum)) %>% 
  slice(1:15)%>% 
  ungroup()

s1 <- gene_set %>% 
  select(GeneSpecies1) %>% 
  filter(grepl("PICAB", GeneSpecies1)) %>% 
  rename("Genes" = GeneSpecies1)
  
s2 <- gene_set %>% 
  select(GeneSpecies2) %>% 
  filter(grepl("PICAB", GeneSpecies2)) %>% 
  rename("Genes" = GeneSpecies2)

gene_set_nor <- rbind(s1,s2)
gene_set_nor <- gene_set_nor %>% 
  distinct(Genes)

# Expressed genes w/ annotation
expressed_genes <- read_delim("data/ComPlEx/transcriptomicsData/NorWood_transcriptomics.txt")
file_1_ori <- read_delim("data/ComPlEx/Picab02_230926_at01_longest_merged.tsv") %>% 
  dplyr::select(id, eggnog_description, interpro_go) 

file_1 <- file_1_ori%>% 
  mutate(id = gsub("\\.\\mRNA\\.\\d+", "",id)) %>% 
  mutate(id = sub("^(.*?_.*?_).*_(.*)$","\\1\\2" , id)) %>% 
  dplyr::rename(ID = "id") 

file_2_and_3 <-inner_join(read_delim("data/ComPlEx/ID_convert_oldtranscript2newGene.txt", col_names = FALSE),            
  read_delim("data/ComPlEx/ID_TRANSLATION_TABLE_clean.txt", col_names = FALSE),by = "X1") %>%
  dplyr::rename(ID = "X1", ID2 = "X2.x", ID3 = "X2.y")

go_list_file <- left_join(file_2_and_3, file_1, by = "ID") %>% 
  dplyr::select(-c(ID, ID3)) %>% 
  separate_rows("interpro_go", sep =  ",") %>% 
  stats::na.omit(interpro_go) %>% 
  dplyr::rename(`Sequence Name` = "ID2", `GO Term` = "eggnog_description", `GO ID` = "interpro_go") %>% 
 filter(`Sequence Name` %in% expressed_genes$Genes)

library(GSEABase)
library(GOstats)
library(gdata)

go_list <- go_list_file

go_data_frame <- data.frame(go_id=go_list$`GO ID`, evidence=rep("ND", nrow(go_list)), 
                          gene_id=go_list$`Sequence Name`)
go_frame_object=GOFrame(go_data_frame,organism="NorwaySpruce")
go_all_frame_object=GOAllFrame(go_frame_object)

gene_set_collection <- GeneSetCollection(go_all_frame_object, setType = GOCollection())
universe <- expressed_genes$Genes

gene_ID <- gene_set_nor$Genes
group_name <- "conserved_genes"
pval_cutoff <- 0.05
params <- GSEAGOHyperGParams(name="GSEA",
                             geneSetCollection = gene_set_collection,
                             geneIds = gene_ID,
                             universeGeneIds = universe, 
                             ontology = "BP",
                             pvalueCutoff = pval_cutoff,
                             conditional = FALSE,
                             testDirection = "over")

hyper_g_test <- hyperGTest(params)
summary_table <- summary(hyper_g_test)
summary_table <- summary_table[summary_table$Count>1,]
summary_table <- summary_table[,c(1,2,5,6,7)]
colnames(summary_table) <- c("GO id","P-value","x","n","GO term")

rownames(summary_table) <- 1:nrow(summary_table)

summary_table$`P-value`<- format(summary_table$`P-value`, digits=3, scientific=TRUE)

# Save
file_for_saving <- paste0(path,"SUMMARY_TBL_", file)
save(summary_table, file = file_for_saving)


```

# Visualising enriched GO terms - treemap
**Step one: load (if needed) summary table - print unique GO IDs and paste in REVIGO**
```{r}
# Print out GO IDs for REVIGO
load("data/ComPlEx/SUMMARY_TBL_clique_GOterms.RData")

df <- summary_table[,1]

for (i in df){
  cat(i, "\n")
}

```

**Step 2: download R script in REVIGO for creating treemap, and copy the revigo.data terms to paste in chunk below**
```{r}
# A treemap R script produced by the Revigo server at http://revigo.irb.hr/
# If you found Revigo useful in your work, please cite the following reference:
# Supek F et al. "REVIGO summarizes and visualizes long lists of Gene Ontology
# terms" PLoS ONE 2011. doi:10.1371/journal.pone.0021800

# author: Anton Kratz <anton.kratz@gmail.com>, RIKEN Omics Science Center, Functional Genomics Technology Team, Japan
# created: Fri, Nov 02, 2012  7:25:52 PM
# last change: Fri, Nov 09, 2012  3:20:01 PM

# -----------------------------------------------------------------------------
# If you don't have the treemap package installed, uncomment the following line:
# install.packages( "treemap" );
library(treemap) 								# treemap package by Martijn Tennekes
library(RColorBrewer)

# Set the working directory if necessary
# # set working directory to the repository root before running, e.g. setwd(".");

# --------------------------------------------------------------------------
# Here is your data from Revigo. Scroll down for plot configuration options.

revigo.names <- c("term_ID","description","frequency","uniqueness","dispensability","representative");
revigo.data <- rbind(c("GO:0006163","purine nucleotide metabolic process",2.3964667573855727,0.8514406248422567,0.06378449,"purine nucleotide metabolic process"),
c("GO:0006396","RNA processing",4.402491558667012,0.7313944163802583,0.6370804,"purine nucleotide metabolic process"),
c("GO:0016070","RNA metabolic process",8.105171248384922,0.8071826712381983,0.52469067,"purine nucleotide metabolic process"),
c("GO:0016072","rRNA metabolic process",1.615169337882767,0.8255055650264516,0.3517316,"purine nucleotide metabolic process"),
c("GO:0006518","peptide metabolic process",0.3993126721131813,0.959558401278688,-0,"peptide metabolic process"),
c("GO:0007018","microtubule-based movement",0.5954751176662081,0.9784174384533229,0,"microtubule-based movement"),
c("GO:0008152","metabolic process",58.08643542977757,1,-0,"metabolic process"),
c("GO:0009058","biosynthetic process",28.507692540195496,0.9201593429544064,0.09765441,"biosynthetic process"),
c("GO:0043170","macromolecule metabolic process",33.33895178165279,0.917149189905993,0.19349584,"biosynthetic process"),
c("GO:0044238","primary metabolic process",48.680252087830326,0.9092797489219028,0.2504376,"biosynthetic process"),
c("GO:0009266","response to temperature stimulus",0.3946804058731623,0.9781143450842387,-0,"response to temperature stimulus"),
c("GO:0043043","peptide biosynthetic process",0.22570852628904622,0.9142618633451162,0.03919999,"peptide biosynthetic process"),
c("GO:0002181","cytoplasmic translation",0.37754692803215667,0.821729355213126,0.40773812,"peptide biosynthetic process"),
c("GO:0006412","translation",4.432288697403288,0.7643879599976175,0.30446211,"peptide biosynthetic process"),
c("GO:0009059","macromolecule biosynthetic process",16.264263355707197,0.7699211293874665,0.54924243,"peptide biosynthetic process"),
c("GO:0010467","gene expression",13.02581206063927,0.7525992895162933,0.59009057,"peptide biosynthetic process"),
c("GO:0018279","protein N-linked glycosylation via asparagine",0.10344574151727937,0.8494382056922374,0.13838005,"peptide biosynthetic process"),
c("GO:0019538","protein metabolic process",18.90831760569825,0.8418949791380846,0.30625735,"peptide biosynthetic process"),
c("GO:0043604","amide biosynthetic process",0.8768225272473001,0.9008698725190911,0.16568804,"peptide biosynthetic process"),
c("GO:0044249","cellular biosynthetic process",23.43500411119782,0.8128193636087967,0.39290026,"peptide biosynthetic process"),
c("GO:0043603","amide metabolic process",1.8259719107448442,0.9511333390477806,0.0478028,"amide metabolic process"),
c("GO:0043632","modification-dependent macromolecule catabolic process",1.2894658869034905,0.8449899419595214,0.05451051,"modification-dependent macromolecule catabolic process"),
c("GO:0051304","chromosome separation",0.03391498221106387,0.9625719852473928,0.0114033,"chromosome separation"),
c("GO:0000070","mitotic sister chromatid segregation",0.1973955833778141,0.7745431490577203,0.69696266,"chromosome separation"),
c("GO:0071840","cellular component organization or biogenesis",14.26840886478907,0.9674642747267268,0.02114748,"cellular component organization or biogenesis"),
c("GO:0072521","purine-containing compound metabolic process",3.3101012792586566,0.9467389271632,0.06144455,"purine-containing compound metabolic process"),
c("GO:0090174","organelle membrane fusion",0.23651632706477607,0.7754830226996999,0.01338594,"organelle membrane fusion"),
c("GO:0000470","maturation of LSU-rRNA",0.1726835998231961,0.6187463340619345,0.69489856,"organelle membrane fusion"),
c("GO:0006996","organelle organization",6.55438598080312,0.7472664488941689,0.61792488,"organelle membrane fusion"),
c("GO:0022613","ribonucleoprotein complex biogenesis",2.67307113984418,0.7245054523789327,0.67345966,"organelle membrane fusion"),
c("GO:0042254","ribosome biogenesis",2.27256471155331,0.6831192197153307,0.62817437,"organelle membrane fusion"),
c("GO:0044085","cellular component biogenesis",6.870847051478853,0.7558214535565586,0.38265648,"organelle membrane fusion"),
c("GO:0048284","organelle fusion",0.2981486047746998,0.7860077729093158,0.42933308,"organelle membrane fusion"),
c("GO:0051276","chromosome organization",1.530548023678806,0.7518370761245676,0.50954628,"organelle membrane fusion"),
c("GO:0070925","organelle assembly",1.5672344893044614,0.7056574015628619,0.61099067,"organelle membrane fusion"),
c("GO:0140694","non-membrane-bounded organelle assembly",0.9606291336269005,0.707878336664542,0.6818075,"organelle membrane fusion"),
c("GO:1990542","mitochondrial transmembrane transport",0.330613850101932,0.8082356167732634,0.0137997,"mitochondrial transmembrane transport"),
c("GO:0006605","protein targeting",0.7521820755319316,0.7146907042460842,0.65382244,"mitochondrial transmembrane transport"),
c("GO:0006839","mitochondrial transport",0.3913575793991847,0.86017658493876,0.2442168,"mitochondrial transmembrane transport"),
c("GO:0033036","macromolecule localization",5.142002589244843,0.8221972403174024,0.30748093,"mitochondrial transmembrane transport"),
c("GO:0033365","protein localization to organelle",1.3430815380758254,0.7057109548028843,0.67571739,"mitochondrial transmembrane transport"),
c("GO:0044743","protein transmembrane import into intracellular organelle",0.1904767202531098,0.7144399708549365,0.68974144,"mitochondrial transmembrane transport"),
c("GO:0051641","cellular localization",5.613085851842994,0.8032148147267274,0.4475155,"mitochondrial transmembrane transport"),
c("GO:0070585","protein localization to mitochondrion",0.23186437000120752,0.7233596030829633,0.22217812,"mitochondrial transmembrane transport"),
c("GO:0071806","protein transmembrane transport",0.5926962501927855,0.7151757546587968,0.55530975,"mitochondrial transmembrane transport"));

stuff <- data.frame(revigo.data);
names(stuff) <- revigo.names;

stuff$frequency <- as.numeric( as.character(stuff$frequency) );
stuff$uniqueness <- as.numeric( as.character(stuff$uniqueness) );
stuff$dispensability <- as.numeric( as.character(stuff$dispensability) );

treemap(
    fontsize.title = 20,
  stuff,
  index = c("representative","description"),
  vSize = "uniqueness",
  type = "categorical",
  vColor = "representative",
  algorithm = "pivotSize",
  inflate.labels = T,      # set this to TRUE for space-filling group labels - good for posters
  lowerbound.cex.labels = 0.4,   # try to draw as many labels as possible (still, some small squares may not get a label) - increase to get fewer
  bg.labels = "#CCCCCC00",   # define background color of group labels
  # "#CCCCCC00" is fully transparent, "#CCCCCCAA" is semi-transparent grey, NA is opaque
  position.legend = "none",
  palette = c("#F7FBFF", "#DEEBF7", "#C6DBEF", "#9ECAE1", "#6BAED6", "#4292C6", "#2171B5"),
  fontsize.labels = c(10,0), # Set last number to 0 to hide children
  border.col = c("black", "#CCCCCC00")
)

```



# Expression profiles

**Step 1: create data frame for looking up GO IDs/terms that actually are associated with genes in gene set (lower/higher order terms)**

```{r}
load("data/ComPlEx/go_list.RData")
load("data/ComPlEx/SUMMARY_TBL_conserved_genes_SUM_15.RData")
load("data/ComPlEx/cliques/conserved_genes_SUM_15coexpr005_full.RData")
#"differentiated_genes_SUM_C-X_3-9.RData"
#"conserved_genes_SUM_15.RData"

gene_set <- conserved_genes %>% # One clique per orthogroup
  ungroup() %>% 
  group_by(OrthoGroup) %>% 
  arrange(desc(NegLog10CliqueSum)) %>% 
  slice(1:15)%>% 
  ungroup()

s1 <- gene_set %>% 
  select(GeneSpecies1, OrthoGroup) %>% 
  rename("Genes" = GeneSpecies1)
  
s2 <- gene_set %>% 
  select(GeneSpecies2, OrthoGroup) %>% 
  rename("Genes" = GeneSpecies2)

gene_set <- rbind(s1,s2)
gene_set <- gene_set %>% 
  distinct(Genes, .keep_all = T)

genes <- gene_set

ortho_general_filtering <- ORTHOLOG_GROUP_FILE %>%
  dplyr::select(Col0 = Arabidopsis_thaliana,
         Aspen = Populus_tremula,
         Birch = Betula_pendula,
         Spruce = Picea_abies,
         Pine = Pinus_sylvestris,
         OrthoGroup = Ortholog_Group) %>% 
  mutate(Ost0 = Col0) %>% 
  dplyr::select(OrthoGroup, all_species)


all_species <- c("Col0", "Ost0", "Aspen","Birch","Spruce","Pine")
file_list_2 <- c()
for (i in all_species){
  expr_name <- paste0("data/expression/",i,"_root_cold_stress_expression.txt")
  file_list_2 <- append(file_list_2, expr_name, after = length(file_list_2))
}

look_up_GO <- go_list %>% 
  filter(`GO ID` %in% summary_table$`GO id`) %>% 
  filter(`Sequence Name` %in% genes$Genes) %>% 
  left_join(summary_table, join_by(`GO ID` == `GO id`)) %>% 
  select(-c(4:7))

length(unique(look_up_GO$`GO ID`))

```

**Step 2: enter the GO ID and run code - gives the expression profiles for all genes associated with the GO ID for all six species**

Note! Check how many genes are present for each GO term. Looking only at max five curves.
```{r}
GO_ID <- "GO:0009409" # For search by gene name set GO_ID = " ". Else:  "GO:0000902"
GO <- "response to cold" 
# OR
gene_name <- "Potra2n5c12117" # Set GO_ID to " "!

palette_1 <- c("#440154FF","black","#35B779FF", "#fc8d59","#FDE725FF")
linethickness <- 1.5

if(!(GO_ID == " ")){
  
  print("Search by GO_ID!")
  
  GO_terms_to_genes <-go_list %>% 
  filter(`Sequence Name` %in% genes$Genes) %>% 
  filter(`GO ID` %in% GO_ID) %>% 
  group_by(`Sequence Name`) %>% 
  slice(1)

genes_to_plot <- unique(GO_terms_to_genes$`Sequence Name`)[1:5]

ortholog_groups <- ortho_general_filtering %>%  
  filter(OrthoGroup %in% genes$OrthoGroup) %>% 
  select(OrthoGroup, all_of(species_list)) %>% 
  separate_rows(Asp, sep = ", ") %>%
  # mutate(Asp = gsub("Poptre_", "", Asp)) %>%
  # mutate(Asp = gsub("\\.\\d\\.p\\d$", "", Asp)) %>% 
  filter(Aspen %in% genes_to_plot) %>% 
  separate_rows(Col0, sep = ", ") %>%
  separate_rows(Ost0, sep = ", ") %>% 
  separate_rows(Birch, sep = ", ") %>% 
  separate_rows(Spruce, sep = ", ") %>% 
  separate_rows(Pine, sep = ", ") %>% 
  filter(Col0 %in% genes$Genes) %>%
  filter(Ost0 %in% genes$Genes) %>%
  filter(Birch %in% genes$Genes) %>%
  filter(Spruce %in% genes$Genes) %>% 
  filter(Pine %in% genes$Genes) %>% 
  distinct(OrthoGroup, .keep_all = T)

length(unique(ortholog_groups$OrthoGroup))
length(unique(ortholog_groups$Asp))

unique(ortholog_groups$Asp)
unique_OG<-c(unique(ortholog_groups$OrthoGroup))

df_species_name_and_sections <- data.frame(
  short = c(rep("Asp",2), rep("Lodge", 3),rep("Birch", 3), rep("Nor", 2), rep("Cher", 2), rep("Scots", 3)),
  long = c(rep("Aspen", 2),  rep("Lodgepole Pine", 3),  rep("Birch",3),rep("Norway Spruce", 2), rep("Cherry",2) ,rep("Scots Pine",3)),
  sections = c(5,12,
               5,12,25,
               5,14,23,
               5,13,
               4,11,
               6,15,27))

expression_all_longer <- data.frame() 
plots <- list()

# For all species
for(x in 1:length(file_list_2)){
  # x <- 6
  
  species <- sapply(strsplit(file_list_2[x], "transcriptomicsData/transcriptomicsForHeatmaps/"), "[",2)
  species <- sapply(strsplit(species, "Wood"), "[", 1)  
  
  # expressologs <- expressologs_sign %>% 
  #   filter(OrthoGroup == og)
  
  filtering_vector <- ortholog_groups %>%
    filter(OrthoGroup %in% unique_OG) %>%
    rename(Species = species)
  
  expression_vector <- read_delim(file_list_2[x], show_col_types = FALSE)  %>% 
    select(Genes, contains("1.")) %>% 
    filter(Genes %in% filtering_vector$Species)
  
  if(nrow(expression_vector) > 0){
    
    gene_order <- unique(filtering_vector$Species)
    
    samples <- colnames(expression_vector)[-1]
    
    expression_vector_t <- t(expression_vector) 
    colnames(expression_vector_t) <- expression_vector_t[1,]
    
    expression_vector_t <- expression_vector_t[-1, ,drop = F] %>% 
      as.data.frame() %>%
      mutate(Samples = samples)
    
    # expression_vector_t <- expression_vector_t[gene_order]
    longer <- expression_vector_t %>% 
      pivot_longer(-Samples,names_to = "Genes", values_to = "Expression") %>% 
      mutate_at("Expression", as.numeric) %>% 
      mutate(Samples = gsub("^(.+?).", "", Samples)) %>% 
      mutate(Genes = factor(Genes, levels =  gene_order))
    
    longer_end <- longer %>% 
      mutate(Species = rep(species, nrow(longer))) %>% 
      group_by(Genes)
    
    expression_all_longer <- rbind(expression_all_longer,longer_end)
    
      cat("Order in cliques: ", gene_order)
  cat("\n")
   cat("Plotting order: ", unique(longer_end$Genes))
  cat("\n")
  }}


expression_all_longer <- expression_all_longer %>% 
  group_by(Genes, Species) 

for(q in 1:length(unique(expression_all_longer$Species))){
  
  # q <- 6
  species <- unique(expression_all_longer$Species)[q]
  
  df <- expression_all_longer %>% 
    filter(Species == species)

  name_long <- df_species_name_and_sections %>% 
    filter(short == species) %>% 
    select(long) %>% 
    slice(1)
  
  name_long_vect <- c(name_long)  
  sub_title <- paste0(name_long_vect)
  
  sections_df <- df_species_name_and_sections %>% 
    filter(short == species)
  
  species_sections <- sections_df$sections
  
  plots[[length(plots)+1]]  <- 
    ggplot(data = df, aes(x = Samples, y = Expression, group = Genes, col = Genes)) +
    geom_line(linewidth = linethickness, show.legend = T) + scale_color_manual(values = palette_1) +
    labs(title = sub_title, y = "Expression (VST)", x = " ") + 
    geom_vline(xintercept = species_sections, linetype = "dashed", color = "black", linewidth = 0.7)   + 
    theme_minimal() +
    theme(axis.ticks.x = element_blank(), axis.text.x.bottom  = element_blank(), legend.text=element_text(size = 10))
}

title <- paste0(GO)


plot_title <- ggdraw() + draw_label(title, fontface = "italic", size = 20, hjust = 0, vjust = 1)
print(plot_grid(plot_title,NULL, plotlist = plots, ncol = 2, nrow = 4))

} else{ 
  
  GO <- " "
  print("Search by gene") 

  ortholog_groups <- ortholog_group_file %>%  
  filter(OrthoGroup %in% genes$OrthoGroup) %>% 
  select(OrthoGroup, all_of(species_list)) %>% 
  separate_rows(Asp, sep = ", ") %>%
  mutate(Asp = gsub("Poptre_", "", Asp)) %>%
  mutate(Asp = gsub("\\.\\d\\.p\\d$", "", Asp)) %>% 
  filter(Asp %in% gene_name) %>% 
  separate_rows(Birch, sep = ", ") %>%
  separate_rows(Cher, sep = ", ") %>% 
  separate_rows(Nor, sep = ", ") %>% 
  separate_rows(Lodge, sep = ", ") %>% 
  separate_rows(Scots, sep = ", ") %>% 
  mutate(Birch = gsub("Betpen_", "", Birch)) %>%
  mutate(Cher = gsub("Prunavi_", "", Cher)) %>%
  mutate(Nor = gsub("\\.p\\d$", "", Nor)) %>%
  mutate(Cher = gsub("\\.p\\d$", "", Cher)) %>% 
  filter(Cher %in% genes$Genes) %>%
  filter(Birch %in% genes$Genes) %>%
  filter(Nor %in% genes$Genes) %>%
  filter(Scots %in% genes$Genes) %>%
  filter(Lodge %in% genes$Genes) %>% 
  distinct(OrthoGroup, .keep_all = T)
  
  length(unique(ortholog_groups$Asp))
  length(unique(ortholog_groups$OrthoGroup))
  
  unique_OG<-c(unique(ortholog_groups$OrthoGroup))

  expression_all_longer <- data.frame() 
  plots <- list()
  
  for(x in 1:length(file_list_2)){
    # x <- 6
    
    species <- sapply(strsplit(file_list_2[x], "transcriptomicsData/transcriptomicsForHeatmaps/"), "[",2)
    species <- sapply(strsplit(species, "Wood"), "[", 1)
    
    # expressologs <- expressologs_sign %>% 
    #   filter(OrthoGroup == og)
    
    filtering_vector <- ortholog_groups %>%
      filter(OrthoGroup %in% unique_OG) %>%
      rename(Species = species)
    
    expression_vector <- read_delim(file_list_2[x], show_col_types = FALSE)  %>% 
      select(Genes, contains("1.")) %>% 
      filter(Genes %in% filtering_vector$Species)
    
    if(nrow(expression_vector) > 0){
      
      gene_order <- unique(filtering_vector$Species)
      
      samples <- colnames(expression_vector)[-1]
      
      expression_vector_t <- t(expression_vector) 
      colnames(expression_vector_t) <- expression_vector_t[1,]
      
      expression_vector_t <- expression_vector_t[-1, ,drop = F] %>% 
        as.data.frame() %>%
        mutate(Samples = samples)
      
      # expression_vector_t <- expression_vector_t[gene_order]
      longer <- expression_vector_t %>% 
        pivot_longer(-Samples,names_to = "Genes", values_to = "Expression") %>% 
        mutate_at("Expression", as.numeric) %>% 
        mutate(Samples = gsub("^(.+?).", "", Samples)) %>% 
        mutate(Genes = factor(Genes, levels =  gene_order))
      
      longer_end <- longer %>% 
        mutate(Species = rep(species, nrow(longer))) %>% 
        group_by(Genes)
      
      expression_all_longer <- rbind(expression_all_longer,longer_end)
      
      cat("Order in cliques: ", gene_order)
      cat("\n")
      cat("Plotting order: ", unique(longer_end$Genes))
      cat("\n")
      }}
  
  expression_all_longer <- expression_all_longer %>% 
    group_by(Genes, Species) 
  
  for(q in 1:length(unique(expression_all_longer$Species))){
    
    # q <- 6
    species <- unique(expression_all_longer$Species)[q]
    
    df <- expression_all_longer %>% 
      filter(Species == species)

    name_long <- df_species_name_and_sections %>% 
      filter(short == species) %>% 
      select(long) %>% 
      slice(1)
    
    name_long_vect <- c(name_long)  
    sub_title <- paste0(name_long_vect, " - ", unique(df$Genes))
    
    sections_df <- df_species_name_and_sections %>% 
      filter(short == species)
    
    species_sections <- sections_df$sections
    
    plots[[length(plots)+1]]  <- 
      ggplot(data = df, aes(x = Samples, y = Expression, group = Genes, col = Genes)) +
      geom_line(linewidth = linethickness, show.legend = F) + scale_color_manual(values = palette_1) +
      labs(title = sub_title, y = "Expression (VST)", x = " ") + 
      geom_vline(xintercept = species_sections, linetype = "dashed", color = "black", linewidth = 0.7) +
      theme_minimal() +
      theme(axis.ticks.x = element_blank(), axis.text.x.bottom  = element_blank(), legend.text=element_text(size = 10))
  }
  
  title <- paste0(GO, " ",unique_OG)
  
  plot_title <- ggdraw() + draw_label(title, fontface = "italic", size = 20, hjust = 0, vjust = 1)
  print(plot_grid(plot_title,NULL, plotlist = plots, ncol = 2, nrow = 4))
  
}



```








```{r}

GO_ID <- c("")
GO <- c("Phloem: ")

unique_OG<-c(unique(ortholog_groups$OrthoGroup))
# palette_1 <- RColorBrewer::brewer.pal(name = "Dark2", n = 8)
# palette_1 <- RColorBrewer::brewer.pal(name = "PRGn", n =11)[c(1:2,8:10)]
palette_1 <- c("#3D5A71","#FFB000","#785EF0", "#DC267F","#FE6100")
linethickness <- 1.5

expression_all_longer <- data.frame() 
plots <- list()

# For all species
for(x in 1:length(file_list_2)){
  # x <- 6
  
  
  species <- sapply(strsplit(file_list_2[x], "transcriptomicsData/transcriptomicsForHeatmaps/"), "[",2)
  species <- sapply(strsplit(species, "Wood"), "[", 1)  
  
  
  
  # expressologs <- expressologs_sign %>% 
  #   filter(OrthoGroup == og)
  
  filtering_vector <- ortholog_groups %>%
    filter(OrthoGroup %in% unique_OG) %>%
    rename(Species = species) 
  
  
  expression_vector <- read_delim(file_list_2[x], show_col_types = FALSE)  %>% 
    select(Genes, contains("1.")) %>% 
    filter(Genes %in% filtering_vector$Species)
  
  if(nrow(expression_vector) > 0){
    
    gene_order <- unique(filtering_vector$Species)
    
    samples <- colnames(expression_vector)[-1]
    
    expression_vector_t <- t(expression_vector) 
    colnames(expression_vector_t) <- expression_vector_t[1,]
    
    expression_vector_t <- expression_vector_t[-1, ,drop = F] %>% 
      as.data.frame() %>%
      mutate(Samples = samples)
    
    # expression_vector_t <- expression_vector_t[gene_order]
    
    
    longer <- expression_vector_t %>% 
      pivot_longer(-Samples,names_to = "Genes", values_to = "Expression") %>% 
      mutate_at("Expression", as.numeric) %>% 
      mutate(Samples = gsub("^(.+?).", "", Samples)) %>% 
      mutate(Genes = factor(Genes, levels =  gene_order))
      
      
    
    
    
    longer_end <- longer %>% 
      mutate(Species = rep(species, nrow(longer))) %>% 
      group_by(Genes)
    

    
    
    expression_all_longer <- rbind(expression_all_longer,longer_end)
    
    
      cat("Order in cliques: ", gene_order)
  cat("\n")
   cat("Plotting order: ", unique(longer_end$Genes))
  cat("\n")
  
  
  } 
  
}


expression_all_longer <- expression_all_longer %>% 
  group_by(Genes, Species) 



for(q in 1:length(unique(expression_all_longer$Species))){
  
  # q <- 6
  species <- unique(expression_all_longer$Species)[q]
  
  
  df <- expression_all_longer %>% 
    filter(Species == species)
  
  
  # if(length(unique(df$Genes))>2){
  #   
  #   linethickness <- 1.5
  #   
  # } else{
  #   linethickness <- 2.5
  # }
  # 
  name_long <- df_species_name_and_sections %>% 
    filter(short == species) %>% 
    select(long) %>% 
    slice(1)
  
  name_long_vect <- c(name_long)  
  sub_title <- paste0(name_long_vect, " - ", unique(df$Genes))
  
  sections_df <- df_species_name_and_sections %>% 
    filter(short == species)
  
  species_sections <- sections_df$sections
  
 
  
  plots[[length(plots)+1]]  <- 
    ggplot(data = df, aes(x = Samples, y = Expression, group = Genes, col = Genes)) +
    geom_line(linewidth = linethickness, show.legend = F) + scale_color_manual(values = palette_1) +
    labs(title = sub_title, y = "Expression (VST)", x = " ") + 
    geom_vline(xintercept = species_sections, linetype = "dashed", color = "black", linewidth = 0.7) +
    theme_minimal() +
    theme(axis.ticks.x = element_blank(), axis.text.x.bottom  = element_blank(), legend.text=element_text(size = 10))
  
  
  
}

title <- paste0(GO, " ",unique_OG)


plot_title <- ggdraw() + draw_label(title, fontface = "italic", size = 20, hjust = 0, vjust = 1)
print(plot_grid(plot_title,NULL, plotlist = plots, ncol = 2, nrow = 4))



```